EGGs are adhesive!\\ An algebraic account of equality saturation

### 2025-09-24 jlamp/jlamp.tex
minori

%String Diagrams
\usepackage{tikz}
\usepackage[draft]{tikzit}
\input{hypergraph.tikzdefs}
\input{hypergraph.tikzstyles}

@@

\subsection{Sharing} 
The classical view of term graphs is that they represent terms as labelled hypergraphs, instead of trees, in order to
exploit the sharing of sub-terms for efficiency reasons. The introduction of equivalence adds another layer of
expressiveness.
%
%Terms are trees, thus different term graphs may represent the same term, up-to the sharing of sub-terms.
%
Consider e.g. a signature $\Sigma$ with a constant $a$ and a binary operator $/$: the term $a / a$ admits a few different 
representations as a term graph with equivalence, such as those below

@@

\begin{center}\begin{tikzpicture}[baseline=(w)]
        \begin{pgfonlayer}{nodelayer}
                \node[style=small box](a) at (-1.8, 0){$a$};
                \node[style=none] (aatt) at (-1.5, 0){};
                \node[style=node](v) at (-0.9, 0){};
                \node[style=medium box](div) at (0, 0){$/$};
                \node[style=none](divfst) at (-0.3, -0.3){};
                \node[style=none] (divsnd) at (-0.3, 0.3){};
                \node[style=none] (divout) at (0.3, 0){};
                \node[style=node] (w) at (0.9, 0){};
        \end{pgfonlayer}
        \begin{pgfonlayer}{edgelayer}
                \draw(aatt.center) to (v);
                \draw[out=-60, in=-180] (v) to (divfst.center);
                \draw[out=60, in=-180](v) to (divsnd.center);
                \draw(divout.center) to (w);
        \end{pgfonlayer}
\end{tikzpicture}
\qquad
\begin{tikzpicture}[baseline=(w)]
        \begin{pgfonlayer}{nodelayer}
                \node[style=small box](a1) at (-1.8, -0.5){$a$};
                \node[style=none] (a1att) at (-1.5, -0.5){};
                \node[style=node](v1) at (-0.9, -0.5){};
                \node[style=small box](a2) at (-1.8, 0.5){$a$};
                \node[style=none] (a2att) at (-1.5, 0.5){};
                \node[style=node](v2) at (-0.9, 0.5){};
                \node[style=medium box](div) at (0, 0){$/$};
                \node[style=none](divfst) at (-0.3, -0.3){};
                \node[style=none] (divsnd) at (-0.3, 0.3){};
                \node[style=none] (divout) at (0.3, 0){};
                \node[style=node] (w) at (0.9, 0){};
        \end{pgfonlayer}
        \begin{pgfonlayer}{edgelayer}
                \draw(a1att.center) to (v1);
                \draw(a2att.center) to (v2);
                \draw[out=0, in=-120] (v1) to (divfst.center);
                \draw[out=0, in=120](v2) to (divsnd.center);
                \draw(divout.center) to (w);
        \end{pgfonlayer}
        \begin{pgfonlayer}{eqlayer}
                \draw[dashed, rounded corners] (-1.2, -0.8) rectangle (-0.6, 0.8);
        \end{pgfonlayer}
\end{tikzpicture}
\qquad
\begin{tikzpicture}[baseline=(w)]
        \begin{pgfonlayer}{nodelayer}
                \node[style=small box](a1) at (-1.8, -0.5){$a$};
                \node[style=none] (a1att) at (-1.5, -0.5){};
                \node[style=node](v1) at (-0.9, -0.5){};
                \node[style=small box](a2) at (-1.8, 0.5){$a$};
                \node[style=none] (a2att) at (-1.5, 0.5){};
                \node[style=node](v2) at (-0.9, 0.5){};
                \node[style=medium box](div) at (0, 0){$/$};
                \node[style=none](divfst) at (-0.3, -0.3){};
                \node[style=none] (divsnd) at (-0.3, 0.3){};
                \node[style=none] (divout) at (0.3, 0){};
                \node[style=node] (w) at (0.9, 0){};
        \end{pgfonlayer}
        \begin{pgfonlayer}{edgelayer}
                \draw(a1att.center) to (v1);
                \draw(a2att.center) to (v2);
                \draw[out=0, in=-120] (v1) to (divfst.center);
                \draw[out=0, in=120](v2) to (divsnd.center);
                \draw(divout.center) to (w);
        \end{pgfonlayer}
\end{tikzpicture}
\end{center}

@@

Consider the term rewriting rule such $x / x \to 1$, from the key example in~\cite{WillseyNWFTP21}.
This is a problematic rule to be simulates using classical term graph rewriting, since it is not left-linear
(that is, $x$ occurs twice in the left-hand side)~\cite{AriolaKP00}.
Using e-term graphs (variables are represented as nodes that do not occur among the targets of an edge),
 it can be modelled as the DPO rule below, concisely given by the arrow $L \rightarrow R$
\[
	\hspace{.25cm}
	\xymatrix{        
		\begin{tikzpicture}[baseline=(w.base)]\begin{pgfonlayer}{nodelayer}
				\node[style=node](v1) at (-0.9, 0.3){};
				\node[style=node](v2) at (-0.9, -0.3){};
				\node[style=medium box] at (0, 0){$/$};
				\node[style=none](divfst) at (-0.3, 0.3){};
				\node[style=none](divsnd) at (-0.3, -0.3){};
				\node[style=none](divout) at (0.3, 0){};
				\node[style=node] (w) at (0.9, 0){};
			\end{pgfonlayer}        
			\begin{pgfonlayer}{edgelayer}
				\draw(v1) to (divfst.center);
				\draw(v2) to (divsnd.center);
				\draw(divout.center) to (w);
			\end{pgfonlayer}
			\begin{pgfonlayer}{eqlayer}
				\draw[dashed, rounded corners](-1.2, 0.6) rectangle (-0.6, -0.6);
			\end{pgfonlayer}\begin{pgfonlayer}{background}
				\draw[color=white] (-1.5, 0.2) rectangle (1.5, -0.2);
			\end{pgfonlayer}
		\end{tikzpicture}
		\ar@{=>}[r] &
		\begin{tikzpicture}[baseline=(nil.center)]\begin{pgfonlayer}{nodelayer}
				\node[style=node](v1) at (-0.9, 0.3){};
				\node[style=node](v2) at (-0.9, -0.3){};
				\node[style=medium box] at (0, 0){$/$};
				\node[style=none](divfst) at (-0.3, 0.3){};
				\node[style=none](divsnd) at (-0.3, -0.3){};
				\node[style=none](divout) at (0.3, 0){};
				\node[style=node] (w) at (0.9, 0){};
				\node[style=small box] at (0, -1){$1$};
				\node[style=none](one) at (0.3, -1){};
				\node[style=node](z) at (0.9, -1){};
				\node[style=none](nil) at (0, -0.5){};
			\end{pgfonlayer}        
			\begin{pgfonlayer}{edgelayer}
				\draw(v1) to (divfst.center);
				\draw(v2) to (divsnd.center);
				\draw(divout.center) to (w);
				\draw(one.center) to (z);
			\end{pgfonlayer}
			\begin{pgfonlayer}{eqlayer}
				\draw[dashed, rounded corners](-1.2, 0.6) rectangle (-0.6, -0.6);
				\draw[dashed, rounded corners](0.6, 0.3) rectangle (1.2, -1.3);
			\end{pgfonlayer}\begin{pgfonlayer}{background}
				\draw[color=white] (-1.5, 0.2) rectangle (1.5, -0.2);
		\end{pgfonlayer}\end{tikzpicture}
	}
\]

In this case $L$ is the minimally-shared e-term graph corresponding to the term $x / x$.
%thus the rule could be applied to any of the term graph with equivalence
%depicted above with a match that is injective on equivalence classes.
The rule cannot be applied to the term graphs with equivalence
 in the pictures (1) and (2) above, while the result of the application of the rule 
to the e-term graphs in the pictures (3) and (4) is shown below

\[xxx\]

@@

The fact that the right-hand side is a regular mono ensures that if the e-hypergraph to whom a rule is applied is acyclic,
the same property holds for the e-hypergraph obtained as the result.
%
Note however that cycles may implicitly occur due to the presence of the equivalence 
on nodes for what are called collapsing rules, i.e. where the right-hand side is a single variable. 
%
Looking again at the key example of~\cite{WillseyNWFTP21}, consider the rule $1 \times x = x$,
which can be modelled as the DPO rule below
%
\[xxx\]
%
Intuitively, the e-term graph obtained after the application of the rule represents a recursive term such 
as $rec_x. 1 \times x$, thus highlighting the difficulty of extracting a term representation out of an e-term graph.

@@

\emph{Related works.}
%\todo{espandere un poco}
Despite the interest they have ben raising as an efficient data structure, we are not aware of any attempt to provide an 
algebraic characterisation of EGGs, the only exception being~\cite{ghica}. We leave for future work an in-depth comparison 
among the two proposals, which appear to be related yet quite different.
%
In fact, for both proposals the key intuition is  to use string diagrams to represents term graphs. We adopted a
more down-to-earth approach by equipping
concretely the nodes of a term graph with an equivalence relation, thus directly corresponding to the original presentation~\cite{DetlefsNS05},
at the same time extending and generalising it~\cite{concur2006} in the contemporary jargon of adhesive categories.
The route chosen in~\cite{ghica} is to consider term graphs as arrows of a symmetric monoidal category,
and equipping them with an enrichment over semi-lattices, the resulting formalism being reminiscent of hierarchical hypergraphs.
Thus, the solution in~\cite{ghica}  is  more general than ours since it can be lifted to term graphs defined over categories other than
$\Set$. At the same time, we both consider the DPO approach for rewriting, even if only our solution guarantees that the resulting category
is actually $\mathcal{M}$-adhesive, thus allowing to exploit all the features of the framework
as they hold for DPO graph transformation.

%\section*{Acknowledgments}
%Hypergraphs, terms graphs, e-graphs, adhesive categories.

>>>>

%String Diagrams
\usepackage{tikz}
\usetikzlibrary{fit}
\usepackage[draft]{tikzit}
\input{hypergraph.tikzdefs}
\input{hypergraph.tikzstyles}

@@

\subsection{Sharing} 
The classical view of term graphs is that they represent terms as labelled hypergraphs, instead of trees, in order to
exploit the sharing of sub-terms for efficiency reasons. The introduction of equivalence adds another layer of
expressiveness.
%
%Terms are trees, thus different term graphs may represent the same term, up-to the sharing of sub-terms.
%
Consider e.g. a signature $\Sigma$ with a constant $a$ and a binary operator $/$: the term $a / a$ admits a few different 
representations as a term graph with equivalence, such as those below.

@@

\begin{center}
\begin{tikzpicture}[baseline=(w)]
        \begin{pgfonlayer}{nodelayer}
                \node[style=small box](a1) at (-1.8, -0.5){$a$};
                \node[style=none] (a1att) at (-1.5, -0.5){};
                \node[style=node](v1) at (-0.9, -0.5){};
                \node[style=small box](a2) at (-1.8, 0.5){$a$};
                \node[style=none] (a2att) at (-1.5, 0.5){};
                \node[style=node](v2) at (-0.9, 0.5){};
                \node[style=medium box](div) at (0, 0){$/$};
                \node[style=none](divfst) at (-0.3, -0.3){};
                \node[style=none] (divsnd) at (-0.3, 0.3){};
                \node[style=none] (divout) at (0.3, 0){};
                \node[style=node] (w) at (0.9, 0){};

                 \node [below=.5cm, align=flush center,text width=2cm] at (v1)
                {
                	Figure $(1)$.
                };
        \end{pgfonlayer}
        \begin{pgfonlayer}{edgelayer}
                \draw(a1att.center) to (v1);
                \draw(a2att.center) to (v2);
                \draw[out=0, in=-120] (v1) to (divfst.center);
                \draw[out=0, in=120](v2) to (divsnd.center);
                \draw(divout.center) to (w);
        \end{pgfonlayer}
\end{tikzpicture}
\quad 
\begin{tikzpicture}[baseline=(w)]
	\begin{pgfonlayer}{nodelayer}
		\node[style=small box](a1) at (-2.3, -0.5){$a$};
		\node[style=none] (a1att) at (-2.3, -0.5){};
		\node[style=node](v1) at (-1.4, -0.5){};
		\node[style=node](w1) at (-0.9, -0.5){};
		\node[style=small box](a2) at (-2.3, 0.5){$a$};
		\node[style=none] (a2att) at (-2.3, 0.5){};
		\node[style=node](v2) at (-1.4, 0.5){};
		\node[style=node](w2) at (-0.9, 0.5){};
		\node[style=medium box](div) at (0, 0){$/$};
		\node[style=none](divfst) at (-0.3, -0.3){};
		\node[style=none] (divsnd) at (-0.3, 0.3){};
		\node[style=none] (divout) at (0.3, 0){};
		\node[style=node] (w) at (0.9, 0){};

		      \node [below=.5cm, align=flush center,text width=2cm] at (-1.15,-0.5)
		{
			Figure $(2)$.
		};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw(a1att.center) to (v1);
		\draw(a2att.center) to (v2);
		\draw[out=0, in=-120] (w1) to (divfst.center);
		\draw[out=0, in=120](w2) to (divsnd.center);
		\draw(divout.center) to (w);
	\end{pgfonlayer}
	\begin{pgfonlayer}{eqlayer}
		\draw[dashed, rounded corners] (-1.7, -0.8) rectangle (-0.6, 0.8);
	\end{pgfonlayer}
\end{tikzpicture}
\quad 
\begin{tikzpicture}[baseline=(w)]
	\begin{pgfonlayer}{nodelayer}
		\node[style=small box](a) at (-1.8, 0){$a$};
		\node[style=none] (aatt) at (-1.5, 0){};
		\node[style=node](v) at (-0.9, 0){};
		\node[style=medium box](div) at (0, 0){$/$};
		\node[style=none](divfst) at (-0.3, -0.3){};
		\node[style=none] (divsnd) at (-0.3, 0.3){};
		\node[style=none] (divout) at (0.3, 0){};
		\node[style=node] (w) at (0.9, 0){};
		      \node [below=.5cm, align=flush center,text width=2cm] at (-0.9,-0.5)
		{
			Figure $(3)$.
		};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw(aatt.center) to (v);
		\draw[out=-60, in=-180] (v) to (divfst.center);
		\draw[out=60, in=-180](v) to (divsnd.center);
		\draw(divout.center) to (w);
	\end{pgfonlayer}
\end{tikzpicture}
\quad 
\begin{tikzpicture}[baseline=(w)]
	\begin{pgfonlayer}{nodelayer}
		\node[style=node](w1) at (-0.9, -0.5){};
		\node[style=small box](a2) at (-2.3, 0){$a$};
		\node[style=none] (a2att) at (-2.3, 0){};
		\node[style=node](v) at (-1.4, 0){};
		\node[style=node](w2) at (-0.9, 0.5){};
		\node[style=medium box](div) at (0, 0){$/$};
		\node[style=none](divfst) at (-0.3, -0.3){};
		\node[style=none] (divsnd) at (-0.3, 0.3){};
		\node[style=none] (divout) at (0.3, 0){};
		\node[style=node] (w) at (0.9, 0){};

		      \node [below=.5cm, align=flush center,text width=2cm] at (-1.15, -0.5)
		{
			Figure $(4)$.
		};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw(a2att.center) to (v);
		\draw[out=0, in=-120] (w1) to (divfst.center);
		\draw[out=0, in=120](w2) to (divsnd.center);
		\draw(divout.center) to (w);
	\end{pgfonlayer}
	\begin{pgfonlayer}{eqlayer}
		\draw[dashed, rounded corners] (-1.7, -0.8) rectangle (-0.6, 0.8);
	\end{pgfonlayer}
\end{tikzpicture}
\end{center}

@@

Consider the term rewriting rule such $x / x \to 1$, from the key example in~\cite{WillseyNWFTP21}.
This is a problematic rule to be simulates using classical term graph rewriting, since it is not left-linear
(that is, $x$ occurs twice in the left-hand side)~\cite{AriolaKP00}.
Using e-term graphs (variables are represented as nodes that do not occur among the targets of an edge),
 it can be modelled as the DPO rule below, concisely given by the arrow $L \rightarrow R$.
\[
	\hspace{.25cm}
	\xymatrix{        
		\begin{tikzpicture}[baseline=(w.base)]\begin{pgfonlayer}{nodelayer}
				\node[style=node](v1) at (-0.9, 0.3){};
				\node[style=node](v2) at (-0.9, -0.3){};
				\node[style=medium box] at (0, 0){$/$};
				\node[style=none](divfst) at (-0.3, 0.3){};
				\node[style=none](divsnd) at (-0.3, -0.3){};
				\node[style=none](divout) at (0.3, 0){};
				\node[style=node] (w) at (0.9, 0){};
			\end{pgfonlayer}        
			\begin{pgfonlayer}{edgelayer}
				\draw(v1) to (divfst.center);
				\draw(v2) to (divsnd.center);
				\draw(divout.center) to (w);
			\end{pgfonlayer}
			\begin{pgfonlayer}{eqlayer}
				\draw[dashed, rounded corners](-1.2, 0.6) rectangle (-0.6, -0.6);
			\end{pgfonlayer}\begin{pgfonlayer}{background}
				\draw[color=white] (-1.5, 0.2) rectangle (1.5, -0.2);
			\end{pgfonlayer}
		\end{tikzpicture}
		\ar@{=>}[r] &
		\begin{tikzpicture}[baseline=(nil.center)]\begin{pgfonlayer}{nodelayer}
				\node[style=node](v1) at (-0.9, 0.3){};
				\node[style=node](v2) at (-0.9, -0.3){};
				\node[style=medium box] at (0, 0){$/$};
				\node[style=none](divfst) at (-0.3, 0.3){};
				\node[style=none](divsnd) at (-0.3, -0.3){};
				\node[style=none](divout) at (0.3, 0){};
				\node[style=node] (w) at (0.9, 0){};
				\node[style=small box] at (0, -1){$1$};
				\node[style=none](one) at (0.3, -1){};
				\node[style=node](z) at (0.9, -1){};
				\node[style=none](nil) at (0, -0.5){};
			\end{pgfonlayer}        
			\begin{pgfonlayer}{edgelayer}
				\draw(v1) to (divfst.center);
				\draw(v2) to (divsnd.center);
				\draw(divout.center) to (w);
				\draw(one.center) to (z);
			\end{pgfonlayer}
			\begin{pgfonlayer}{eqlayer}
				\draw[dashed, rounded corners](-1.2, 0.6) rectangle (-0.6, -0.6);
				\draw[dashed, rounded corners](0.6, 0.3) rectangle (1.2, -1.3);
			\end{pgfonlayer}\begin{pgfonlayer}{background}
				\draw[color=white] (-1.5, 0.2) rectangle (1.5, -0.2);
		\end{pgfonlayer}\end{tikzpicture}
	}
\]

In this case $L$ is the minimally-shared e-term graph corresponding to the term $x / x$.
%thus the rule could be applied to any of the term graph with equivalence
%depicted above with a match that is injective on equivalence classes.
The rule cannot be applied to the term graphs with equivalence
 in the pictures (1) and (2) above, while the result of the application of the rule 
to the e-term graphs in the pictures (3) and (4) is shown below.
\[
	\begin{tikzpicture}[baseline=(w)]
		\begin{pgfonlayer}{nodelayer}
			\node[style=small box](a) at (-1.8, 0){$a$};
			\node[style=none] (aatt) at (-1.5, 0){};
			\node[style=node](v) at (-0.9, 0){};
			\node[style=medium box](div) at (0, 0){$/$};
			\node[style=none](divfst) at (-0.3, -0.3){};
			\node[style=none] (divsnd) at (-0.3, 0.3){};
			\node[style=none] (divout) at (0.3, 0){};
			\node[style=node] (w) at (0.9, 0){};
				\node[style=small box] at (0, -1){$1$};
			\node[style=none](one) at (0.3, -1){};
			\node[style=node](z) at (0.9, -1){};
		\end{pgfonlayer}
		\begin{pgfonlayer}{edgelayer}
			\draw(aatt.center) to (v);
			\draw[out=-60, in=-180] (v) to (divfst.center);
			\draw[out=60, in=-180](v) to (divsnd.center);
			\draw(divout.center) to (w);
			\draw(one.center) to (z);
		\end{pgfonlayer}
		\begin{pgfonlayer}{eqlayer}
			\draw[dashed, rounded corners](0.6, 0.3) rectangle (1.2, -1.3);
		\end{pgfonlayer}
	\end{tikzpicture}
	\qquad \qquad 
	\begin{tikzpicture}[baseline=(w)]
		\begin{pgfonlayer}{nodelayer}
			\node[style=node](w1) at (-0.9, -0.5){};
			\node[style=small box](a2) at (-2.3, 0){$a$};
			\node[style=none] (a2att) at (-2.3, 0){};
			\node[style=node](v) at (-1.4, 0){};
			\node[style=node](w2) at (-0.9, 0.5){};
			\node[style=medium box](div) at (0, 0){$/$};
			\node[style=none](divfst) at (-0.3, -0.3){};
			\node[style=none] (divsnd) at (-0.3, 0.3){};
			\node[style=none] (divout) at (0.3, 0){};
			\node[style=node] (w) at (0.9, 0){};
			\node[style=small box] at (0, -1){$1$};
			\node[style=none](one) at (0.3, -1){};
			\node[style=node](z) at (0.9, -1){};
		\end{pgfonlayer}
		\begin{pgfonlayer}{edgelayer}
			\draw(a2att.center) to (v);
			\draw[out=0, in=-120] (w1) to (divfst.center);
			\draw[out=0, in=120](w2) to (divsnd.center);
			\draw(divout.center) to (w);
			\draw(one.center) to (z);
		\end{pgfonlayer}
		\begin{pgfonlayer}{eqlayer}
			\draw[dashed, rounded corners] (-1.7, -0.8) rectangle (-0.6, 0.8);
			\draw[dashed, rounded corners](0.6, 0.3) rectangle (1.2, -1.3);
		\end{pgfonlayer}
	\end{tikzpicture}
\]

@@

The fact that the right-hand side is a regular mono ensures that if the e-hypergraph to whom a rule is applied is acyclic,
the same property holds for the e-hypergraph obtained as the result.
%
Note however that cycles may implicitly occur due to the presence of the equivalence 
on nodes for what are called collapsing rules, i.e. where the right-hand side is a single variable. 
%
Looking again at the key example of~\cite{WillseyNWFTP21}, consider the rule $1 \times x = x$,
which can be modelled as the DPO rule below.
\[
\hspace{.25cm}
\xymatrix{        
	\begin{tikzpicture}[baseline=(w.base)]\begin{pgfonlayer}{nodelayer}
			\node[style=node](v1) at (-0.9, 0.3){};
			\node[style=node](v2) at (-0.9, -0.3){};
			\node[style=medium box] at (0, 0){$\times$};
			\node[style=none](divfst) at (-0.3, 0.3){};
			\node[style=none](divsnd) at (-0.3, -0.3){};
			\node[style=none](divout) at (0.3, 0){};
			\node[style=node] (w) at (0.9, 0){};
			\node[style=small box] at (-2.3, 0.3){$1$};
			\node[style=none](one) at (-2, 0.3){};
			\node[style=node](z) at (-1.4, 0.3){};
		\end{pgfonlayer}        
		\begin{pgfonlayer}{edgelayer}
			\draw(v1) to (divfst.center);
			\draw(v2) to (divsnd.center);
			\draw(divout.center) to (w);
			\draw(one.center) to (z);
		\end{pgfonlayer}
		\begin{pgfonlayer}{eqlayer}
			\draw[dashed, rounded corners](-1.7, 0.6) rectangle (-0.6, -0);
		\end{pgfonlayer}\begin{pgfonlayer}{background}
			\draw[color=white] (-1.5, 0.2) rectangle (1.5, -0.2);
		\end{pgfonlayer}
	\end{tikzpicture}
	\ar@{=>}[r] &	\hspace{10pt}
	\begin{tikzpicture}[baseline=(w.base)]\begin{pgfonlayer}{nodelayer}
			\node[style=node](v1) at (-0.9, 0.3){};
			\node[style=node](v2) at (-0.9, -0.3){};
			\node[style=medium box] at (0, 0){$\times$};
			\node[style=none](divfst) at (-0.3, 0.3){};
			\node[style=none](divsnd) at (-0.3, -0.3){};
			\node[style=none](divout) at (0.3, 0){};
			\node[style=node] (w) at (0.9, 0){};
			\node[style=small box] at (-2.3, 0.3){$1$};
			\node[style=none](one) at (-2, 0.3){};
			\node[style=node](z) at (-1.4, 0.3){};
		\end{pgfonlayer}        
		\begin{pgfonlayer}{edgelayer}
			\draw(v1) to (divfst.center);
			\draw(v2) to (divsnd.center);
			\draw(divout.center) to (w);
			\draw(one.center) to (z);
		\end{pgfonlayer}
		\begin{pgfonlayer}{eqlayer}
			\draw[dashed, rounded corners](-1.7, 0.6) rectangle (-0.6, -0);
		\end{pgfonlayer}
		\begin{pgfonlayer}{background}
			\draw[color=white] (-1.5, 0.2) rectangle (1.5, -0.2);
		\end{pgfonlayer}
		\draw[dashed,rounded corners]
		([shift={(0,0.2)}]v2.north) --
		([shift={(-0.2,0)}]v2.west) --
		(-.9,-1.2)--(.9,-1.2)--([shift={(0.2,0)}]w.east)--
		([shift={(0,0.2)}]w.north)  -- ([shift={(-0.2,0)}]w.west)-- (0.7,-0.8)--(-0.7,-0.8)--([shift={(0.2,0)}]v2.east)-- cycle;
	\end{tikzpicture}
}
\]

Intuitively, the e-term graph obtained after the application of the rule represents a recursive term such 
as $rec_x. 1 \times x$, thus highlighting the difficulty of extracting a term representation out of an e-term graph.

@@

\emph{Related works.}
%\todo{espandere un poco}
Despite the interest they have ben raising as an efficient data structure, we are not aware of any attempt to provide an 
algebraic characterisation of EGGs, the only exception being~\cite{ghica}. We leave for future work an in-depth comparison 
among the two proposals, which appear to be related yet quite different.
%
In fact, for both proposals the key intuition is  to use string diagrams to represents term graphs. We adopted a
more down-to-earth approach by equipping
concretely the nodes of a term graph with an equivalence relation, thus directly corresponding to the original presentation~\cite{DetlefsNS05},
at the same time extending and generalising it~\cite{concur2006} in the contemporary jargon of adhesive categories. 
The route chosen in~\cite{ghica} is to consider term graphs as arrows of a symmetric monoidal category,
and equipping them with an enrichment over semi-lattices, the resulting formalism being reminiscent of hierarchical hypergraphs.
Thus, the solution in~\cite{ghica}  is  more general than ours since it can be lifted to term graphs defined over categories other than
$\Set$. At the same time, we both consider the DPO approach for rewriting, even if only our solution guarantees that the resulting category
is actually $\mathcal{M}$-adhesive, thus allowing to exploit all the features of the framework
as they hold for DPO graph transformation.  Moreover, our construction is technically simpler, which not only facilitates its practical application and theoretical analysis, but also aligns with the traditional methodology of endowing certain (hyper)graphs with an additional \emph{structure}.
%\section*{Acknowledgments}
%Hypergraphs, terms graphs, e-graphs, adhesive categories.
